---
title: "Untitled"
author: "Nathaniel Decker"
date: "11/6/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning = FALSE)
rm(list=ls())
x <- c("tidyverse", 'lubridate', "pander", "geosphere", "mapsapi", 'censusapi', "knitr") 
lapply(x, library, character.only = TRUE)
ci_calc <- function(z, y, n){ # This is the Agresti adjustment
  z * sqrt((((y + 2)/(n+4))*(1-((y + 2)/(n+4))))/(n + 4))
}
censuskey <- 'f8b8ef413394e43d536ae3829d873bf0c246b1ba'
srpos_link <- "../Data/MySurvey/srpos.csv"
srpos <- read_csv(srpos_link) %>% 
  mutate(PROPERTYADDRESSZIP = as.character(PROPERTYADDRESSZIP))
```


## Introduction

Stratify all dagt by portfolion size bins


##  Who Responded to the Survey?

Stratification by metro?

```{r, echo=FALSE, results='asis'}
# I'm not sure this is useful - might be better to just have a map.
srpos %>% dplyr::count(MSANAME) %>% arrange(desc(n)) %>% transmute(`Metro Area` = MSANAME,Responses = n) %>% kable(caption = "Responses by metro of rental property")
```

Might want to include metro of owner, too (though I'll have to geocode that)

Stratify by portfolio size

```{r, echo=FALSE,results='asis'}
srpos %>% mutate(Port_bin = cut(PortfolioEst, c(0,2,10,50, Inf))) %>% dplyr::count(Port_bin) %>% transmute(`Portfolio Size of Owner` = recode(Port_bin,"(0,2]" = "1-2 units","(2,10]" = "3-10 units","(10,50]" = "11-50 units","(50,Inf]" = ">50 units"), Respondents = n) %>% kable(caption="Most Respondents had more than 2, but less than 50 units")
```

##  What are their properties like?

```{r}
srpos %>% mutate(TYPE = factor(recode(TYPE, "Co-operativeunit above" = "Co-operative",
                               "Condominiumunit above" = "Condominium",
                               "Mobile homeunit above" = "Mobile home",
                               "Single family attached house, rowhouse or townhouse (not a condominium)property above" = "Single family attached house, rowhouse or townhouse",
                               "Single family detached houseproperty above" = "Single family detached house",
                               "Unit in a property with 2 to 4 housing unitsproperty containing the rental unit identified above" = "Unit in a property with 2 to 4 housing units",
                               "Unit in a property with 5 or more housing unitsproperty containing the rental unit identified above" = "Unit in a property with 5 or more housing units"), levels = rev(c("Single family detached house","Condominium", "Unit in a property with 2 to 4 housing units","Single family attached house, rowhouse or townhouse", "Unit in a property with 5 or more housing units", "Mobile home","Co-operative")))) %>% drop_na(TYPE) %>% ggplot(aes(x = reorder(TYPE, TYPE))) + geom_bar() + coord_flip()
```

## What are their tenant like?

## Rent, Rent Regulation

## Rent Setting

## Profitiblity & Market

## Management

## Maintenace and Capital Improvments

## Tenant Screening and Marketing

## Turnover

## Section 8

## Tenant Problems

## Local Policies

## Acquisition

# Pland for Ownership

## Financing

## Ownership strucutre

## Owners (push to top)

